### 2018-05-21 paper.tex
initial creation

\documentclass[11pt]{article}

 \pdfminorversion=4

\usepackage{mkolar_definitions}
\usepackage{color}
\usepackage{fullpage}
\usepackage{wrapfig}
\usepackage{algorithm}
\usepackage{algorithmic}
\usepackage{multirow}
\usepackage{tabularx}
\usepackage{array}
\usepackage{natbib}
\usepackage{mathabx}
\usepackage{graphicx}
\usepackage{rotating}
\usepackage{multirow}

\usepackage[usenames,dvipsnames,svgnames,table]{xcolor}
\usepackage[colorlinks=true,
            linkcolor=blue,
            urlcolor=blue,
            citecolor=blue]{hyperref}

\numberwithin{equation}{section}
\numberwithin{theorem}{section}
%\numberwithin{lemma}{section}

%\addtolength{\textwidth}{1in} \addtolength{\oddsidemargin}{-0.5in}
%\addtolength{\textheight}{1in} \addtolength{\topmargin}{-0.62in}
\renewcommand{\baselinestretch}{1.05}

\usepackage{xargs}    
\usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}
\newcommandx{\unsure}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{#2}}
\newcommandx{\change}[2][1=]{\todo[linecolor=blue,backgroundcolor=blue!25,bordercolor=blue,#1]{#2}}
\newcommandx{\info}[2][1=]{\todo[linecolor=OliveGreen,backgroundcolor=OliveGreen!25,bordercolor=OliveGreen,#1]{#2}}
\newcommandx{\improvement}[2][1=]{\todo[linecolor=Plum,backgroundcolor=Plum!25,bordercolor=Plum,#1]{#2}}

\newcommand{\mk}[1]{\textcolor{green}{Mladen: #1}}
\newcommand{\UPDATE}{\fbox{Update}\\}

\begin{document}

\title{}
\author{} 
\date{First draft: ...}

\maketitle

\begin{abstract}

\end{abstract}

\noindent {\bf Keywords:}

\newpage

\section*{Acknowledgments}

This work is partially supported by an IBM Corporation Faculty
Research Fund at the University of Chicago Booth School of Business.
This work was completed in part with resources provided by the
University of Chicago Research Computing Center.

%\newpage

{
\bibliographystyle{my-plainnat}
\bibpunct{(}{)}{,}{a}{,}{,}
\bibliography{paper}
}

%\newpage

\appendix

\end{document}